---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# powRICLPM

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/powRICLPM)](https://CRAN.R-project.org/package=powRICLPM)
[![R-CMD-check](https://github.com/JeroenDMulder/powRICLPM/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JeroenDMulder/powRICLPM/actions/workflows/R-CMD-check.yaml)
 <a href="https://www.r-pkg.org/pkg/powerly"><img src="https://cranlogs.r-pkg.org/badges/grand-total/powerly" alt="CRAN RStudio mirror downloads"/></a>
<!-- badges: end -->


`powRICLPM` is an `R` package that performs a power analysis for the random intercept cross-lagged panel model (RI-CLPM) in a simple and user-friendly way. It implements the strategy as proposed by [Mulder (2022)](https://doi.org/10.1080/10705511.2022.2122467). Its main functionalities include:  

- [Setting up and performing a basic power analysis](https://jeroendmulder.github.io/powRICLPM/articles/start.html): Obtain the power to reject the null-hypothesis of no effect (as well as other performance measures, such as bias, mean square error, etc.) for all parameters in the RI-CLPM given a specific sample size, number of repeated measures, and proportion of between-unit variance (among other things). The power analysis can be performed across multiple experimental conditions simultaneously (i.e., with varying numbers of repeated measures, proportions of between-unit variance, etc.).
- [Extending the basic power analysis setup](https://jeroendmulder.github.io/powRICLPM/articles/extensions.html): Extend the basic power analysis to include the use of bounded estimation, various (stationarity) constraints over time on parameters of the estimation model, and/or the estimation of measurement error. 
- [Create Mplus model syntax](https://jeroendmulder.github.io/powRICLPM/articles/mplus.html): Create syntax for performing RI-CLPM power analyses using Mplus.  

## Documentation

There are four sources of documentation for `powRICLPM`:

- The rationale for the power analysis strategy underlying this package can be found in [Mulder (2022)](https://doi.org/10.1080/10705511.2022.2122467).
- Every user-facing function in the package is documented, and the documentation can be accessed by running `?function_name` in the R console (e.g., `?powRICLPM`). Here, you can find explanations on how to use the functions, as well as technical details.
- There are four main vignettes accessible via the ‘Vignettes’ tab, describing functionalities and analysis options of this package more generally. The ‘Example’ vignette serves as the online supplementary material to [Mulder (2022)](https://doi.org/10.1080/10705511.2022.2122467), and contains the R code for an illustrative example using the `powRICLPM` package.
- The [FAQ](https://jeroendmulder.github.io/powRICLPM/articles/FAQ.html) contains answers to frequently asked question that reach me via email.

## Installation

To install the development version of `powRICLPM`, including the latest bug fixes and new features, run:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("jeroendmulder/powRICLPM")
```

To install the latest release of `powRICLPM` from CRAN, run:

```{r gh-install-CRAN, eval = FALSE}
install.packages("powRICLPM")
```


## Citing `powRICLPM`

You can cite the R-package with the following citation:

> Mulder, J.D., (2022). *Power analysis for the random intercept cross-lagged panel model using the powRICLPM R-package*. Structural Equation Modeling: A Multidisciplinary Journal. https://doi.org/10.1080/10705511.2022.2122467 

## Contact

If you have ideas, comments, or issues you would like to raise, please get in touch.  

* Issues and ideas can be raised on GitHub via [https://github.com/jeroendmulder/powRICLPM](https://github.com/jeroendmulder/powRICLPM)
* Pull request can be created on GitHub via [https://github.com/jeroendmulder/powRICLPM/pulls](https://github.com/jeroendmulder/powRICLPM/pulls)
