---
output: github_document
bibliography: "./vignettes/references.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# powRICLPM

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/powRICLPM)](https://CRAN.R-project.org/package=powRICLPM)
[![R-CMD-check](https://github.com/JeroenDMulder/powRICLPM/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JeroenDMulder/powRICLPM/actions/workflows/R-CMD-check.yaml)
 <a href="https://www.r-pkg.org/pkg/powRICLPM"><img src="https://cranlogs.r-pkg.org/badges/grand-total/powRICLPM" alt="CRAN RStudio mirror downloads"/></a>
<!-- badges: end -->


`powRICLPM` is an R package that aids researchers with performing a power analysis for the random intercept cross-lagged panel model (RI-CLPM) by @hamaker_critique_2015, and the Stable Trait Autoregressive Trait State Model (STARTS) by @kenny_trait-state-error_1995 and @kenny_trait-state_2001. It implements the strategy as proposed by @mulder_power_2023. Its main functionalities include:  

- [**Basic power analysis**](https://jeroendmulder.github.io/powRICLPM/articles/start.html): Use Monte Carlo simulations to compute the power to reject the null-hypothesis (as well as other performance measures such as bias, mean square error) for all parameters in the RI-CLPM and STARTS, for a specific experimental condition. A condition is defined by its sample size, number of repeated measures, proportion of between-unit variance, and reliability of the indicators. `powRICLPM` can perform power analyses across multiple experimental conditions simultaneously, and report the results back in a user-friendly manner.
- [**Extensions**](https://jeroendmulder.github.io/powRICLPM/articles/extensions.html): The basic power analysis setup can be extended to include the use of bounded estimation, various (stationarity) constraints over time on parameters of the estimation model, the generation of nonnormal data, among other things.
- [**Mplus**](https://jeroendmulder.github.io/powRICLPM/articles/mplus.html): When Mplus is installed, `powRICLPM` can create Mplus syntax, and run the power analyses in Mplus. 

## Documentation

There are four sources of documentation for `powRICLPM`:

- The rationale for the power analysis strategy underlying this package can be found in @mulder_power_2023.
- Every user-facing function in the package is documented, and the documentation can be accessed by running `?function_name` in the R console (e.g., `?powRICLPM`). Here, you can find explanations on how to use the functions, as well as technical details.
- More elaborate descriptions of this package's functionality and analysis options are described in vignettes. These are accessible via the 'Vignettes' tab in the menu, or via R using `vignette(package = "powRICLPM")`. 
- The [FAQ](https://jeroendmulder.github.io/powRICLPM/articles/FAQ.html) contains answers to frequently asked question that reach me via email.

## Installation

To install the development version of `powRICLPM`, including the latest bug fixes and new features, run:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("jeroendmulder/powRICLPM")
```

To install the latest release of `powRICLPM` from CRAN, run:

```{r gh-install-CRAN, eval = FALSE}
install.packages("powRICLPM")
```


## Citing `powRICLPM`

You can cite the R-package with the following citation:

> Mulder, J.D., (2023). Power analysis for the random intercept cross-lagged panel model using the powRICLPM R-package. *Structural Equation Modeling: A Multidisciplinary Journal, 30*(4), 645-658. https://doi.org/10.1080/10705511.2022.2122467 

## Contact

If you have ideas, comments, or issues you would like to raise, please get in touch.  

* Issues and ideas can be raised on GitHub via [https://github.com/jeroendmulder/powRICLPM](https://github.com/jeroendmulder/powRICLPM)
* Pull request can be created on GitHub via [https://github.com/jeroendmulder/powRICLPM/pulls](https://github.com/jeroendmulder/powRICLPM/pulls)

## References
