---
title: "Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, eval = FALSE, echo=FALSE}
library(powRICLPM)
library(ggplot2)
library(furrr)
library(progressr)
```

This vignette is a supplement to "Research design planning for the random intercept cross-lagged panel model" by Mulder (under review). It contains the R-code used for performing a power analysis (using the `powRICLPM` package) for the illustrative example. In total there are 171 experimental conditions (19 sample sizes $\times$ 3 numbers of repeated measures $\times$ 3 proportions of between person variance). The analysis is partitioned into two parts to reduce total computation time: A preliminary analysis part, and a validation part. 

## Preliminary analysis
First, a Monte Carlo power analysis is performed over all 171 experimental conditions using a limited number of replications (`reps = 100`). These preliminary results serve as a basis for selecting those experimental conditions that show promising results (i.e., that meet the desired power- and accuracy levels). These conditions are then validated using a large number of replications (`reps = 2000`) in the next step. 

The R-code for the preliminary analysis can be found below:
```{r preliminary-analysis, eval = FALSE}
# Matrix of standardized lagged effects
Phi <- matrix(c(0.20, 0.10, 0.15, 0.30), byrow = FALSE, ncol = 2)

# Correlation matrix of within-components
wSigma <- matrix(c(1, 0.26, 0.26, 1), byrow = FALSE, ncol = 2)

# powRICLPM automatically computes Psi based on Phi and wSigma

# Setup parallel processing to speed up computations
plan(multisession, workers = 6)

# Perform preliminary power analysis (with progress bar)
with_progress({
  out_preliminary <- powRICLPM(target_power = 0.8,
                               search_lower = 200,
                               search_upper = 2000,
                               search_step = 100,
                               time_points = c(3, 4, 5),
                               ICC = c(0.50, 0.55, 0.60),
                               RI_cor = 0.35, 
                               Phi = Phi,
                               wSigma = wSigma,
                               reps = 5, 
                               seed = 20220209)
})

# Tabular summary of results
summary(out_preliminary)
coef_wB2wA1 <- coef_powRICLPM(out_preliminary, parameter = "wB2~wA1")
coef_wB2wA1 # Can be used to check if any issues occurred with estimation 

# Visualize power
p <- plot_powRICLPM(object = out_preliminary,
                    x = "sample_size",
                    y = "pwr",
                    color = "factor(time_points)",
                    wrap = "ICC",
                    parameter = "wB2~wA1")

# Tailor visualization for Mulder (2022)
p <- p + 
  labs(color = "Number of time points") +
  scale_x_continuous(name = "Sample size",
                     breaks = seq(200, 2000, 200),
                     guide = guide_axis(n.dodge = 2)) +
  theme(legend.position = "bottom")
p
#ggsave("Mulder2022_preliminary_power.png", height = 5, width = 7)

# Visualize accuracy
a <- plot_powRICLPM(out_preliminary,
                    "sample_size",
                    "acc",
                    color = "factor(time_points)",
                    wrap = "ICC", 
                    parameter = "wB2~wA1")
a <- a +
  labs(color = "Number of time points") +
  ylab("Accuracy") +
  scale_x_continuous(name = "Sample size",
                     breaks = seq(200, 2000, 200),
                     guide = guide_axis(n.dodge = 2)) +
  theme(legend.position = "bottom")
a
#ggsave("Mulder2022_preliminary_accuracy.png", height = 5, width = 7)
  
```

## Validation
The preliminary results suggest three minimum sample sizes that meet the desired power level: 

- 600 (for conditions with 4 and 5 times points, across all proportions of between-unit variance),
- 1100 (for conditions with 3 repeated measures and with a 0.50 or 0.55 proportion of between-unit variance), and 
- 1200 (for conditions with 3 repeated measures and with a 0.60 proportion of between-unit variance). 

Therefore, power analysis results are validated for a total of 9 experimental conditions using `reps = 1000`. 

```{r validation, eval = FALSE}
# Setup parallel processing to speed up computations
plan(multisession, workers = 5)

# Perform preliminary power analysis (with progress bar)
with_progress({
  out_validation <- powRICLPM(target_power = 0.8,
                              search_lower = 1200,
                              search_upper = 2000,
                              search_step = 100,
                              time_points = c(4, 5),
                              ICC = c(0.50, 0.55, 0.60),
                              RI_cor = 0.35, 
                              Phi = Phi,
                              wSigma = wSigma,
                              reps = 2000, 
                              seed = 20220209)
})

# Tabular summary of results
summary(out_validation, parameter = "wB2~wA1")
coef_wB2wA1 <- coef_powRICLPM(out_validation, parameter = "wB2~wA1")
coef_wB2wA1

# Visualize power
p2 <- plot_powRICLPM(out_validation,
                    "sample_size",
                    "pwr",
                    color = "factor(time_points)",
                    wrap = "ICC",
                    parameter = "wB2~wA1")

# Tailor visualization of power for Mulder (2022)
p2 <- p2 + 
  labs(color = "Number of time points") +
  scale_x_continuous(name = "Sample size",
                     breaks = seq(1200, 2000, 100),
                     guide = guide_axis(n.dodge = 2)) +
  scale_color_manual(values = c("#00BA38", "#619CFF")) +
  theme(legend.position = "bottom")
p2
#ggsave("Mulder2022_validation_power.png", height = 5, width = 7)

# Visualize accuracy
a2 <- plot_powRICLPM(out_validation,
                    "sample_size",
                    "acc",
                    color = "factor(time_points)",
                    wrap = "ICC", 
                    parameter = "wB2~wA1")
a2 <- a2 +
  labs(color = "Number of time points") +
  scale_x_continuous(name = "Sample size",
                     breaks = seq(1200, 2000, 100),
                     guide = guide_axis(n.dodge = 2)) +
  theme(legend.position = "bottom")
a2
#ggsave("Mulder2022_validation_accuracy.png", height = 5, width = 7)

```



