---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette details how you can set up and execute a basic power analysis for a bivariate random intercept cross-lagged panel model (RI-CLPM) using the `powRICLPM` package. Throughout, an illustrating example will be used in which we wish to detect a small cross-lagged effect $\beta_{2}$ (defined here as the effect of $a_{1}^{*}$ to $b_{2}^{*}$, where $a_{1}^{*}$ and $b_{2}^{*}$ denote the latent within-unit components of $a_{1}$ and $b_{2}$, respectively) of 0.2 (standardized). For the design of our power analysis we follow the strategy as described in Mulder (under review). The basic power analysis described here can be extended with various options, as described in the Vignette [Extensions](https://jeroendmulder.github.io/powRICLPM/articles/extensions.html). 

## Step 1: Experimental conditions
Before performing the power analysis, you must first determine the *experimental conditions* of interest. Experimental conditions (or: simulation conditions) are defined by characteristics of the study design that can impact statistical power. This includes characteristics like the *sample size* and the *number of repeated measures*. You must decide on the number of repeated measures that will be used in the simulations, as well as the range of sample sizes over which you want to simulate the power. 

For this example, we take a sample size range from 100 to 1000 first, increasing with steps of 100. We let the numbers of repeated measures range from 3 to 5. If these experimental conditions do not lead to the desired amount of power for detecting the small cross-lagged effect, the ranges can be extended. 

## Step 2: Choosing population parameter values
Next, you must determine population parameter values for generating data from the RI-CLPM. This requires the specification of:

- `Phi`: Standardized autoregressive and cross-lagged effects for the within-unit components of the model. These values are collected in a matrix with columns representing predictors and rows representing outcomes.
- `within_cor`: A correlation for the within-unit components.
- `ICC`: The proportion of variance at the between-unit level (relative to the total variance).
- `RI_cor`: The correlation between the random intercepts.

For our example, the parameter values are set to:

```{r step2}
Phi <- matrix(c(.4, .1, .2, .3), ncol = 2, byrow = T) # The .2 refers to our standardized cross-lagged effect of interest
ICC <- 0.5
RI_cor <- 0.3
```

## Steps 3-5: Performing the power analysis
Steps 3 to 5 are automated by the `powRICLPM` function. As input, you must provide:

- the desired power level using the `target_power` argument, 
- the range of sample sizes to simulate the power for using the `search_lower`, `search_upper`, and `search_step` arguments (alternatively, you can specify this directly by providing a vector of sample sizes to the `sample_size` argument), 
- the number of time points for the simulated data using the `time_points` argument, 
- the population values `Phi`, `within_cor`, `ICC`, and `RI_cor`, and
- the number of Monte Carlo replications we want to perform per experimental condition in the `reps` argument. 

Optionally, you can specify:

- `skewness` and `kurtosis`: An integer (vector) that determines the skewness and kurtosis for the simulated observed variables, respectively. Suppose we have reason to believe the $A$ and $B$ variables are positively skewed, and have heavy tails (i.e., a higher kurtosis) we can include the arguments `skewness = 1` and `kurtosis = 0.5` (defaults: 0).
- `alpha`: A numeric value denoting the significance criterion (default: 0.05).
- `seed`: An integer to control the starting point of the random number generator. This is important to use if you want to replicate the results.
- `reliability`: A numeric value representing the reliability of the indicators (i.e., the proportion of true score variance) (default: 1)

The `constraints`, `bounds`, and `est_ME` arguments can be set as well to extend the basic power analysis setup, as described in the Vignette [Extensions](https://jeroendmulder.github.io/powRICLPM/articles/extensions.html). 

For our example, we can perform the power analysis by running:

```{r analysis, eval = F}
output <- powRICLPM(target_power = 0.8,
                    search_lower = 100,
                    search_upper = 1000,
                    search_step = 50,
                    time_points = c(3, 4, 5),
                    ICC = ICC,
                    RI_cor = RI_cor, 
                    Phi = Phi,
                    within_cor = 0.3,
                    reps = 1000)
```

### Parallel processing using `furrr`
Performing a Monte Carlo power analysis with a large number of replications, and across multiple simulation conditions (in our illustrative example, we have 57 experimental conditions) can take a lot of time. To speed up the process, it is recommended to perform the power analysis *across simulation conditions* in parallel (i.e., on multiple cores). The `powRICLPM()` function has implemented `future`’s parallel processing capabilities using the `furrr` package. 

Load the `furrr` package, and use its `plan()` function to change the power analysis execution from *sequential* (i.e., single-core, the default), to *multisession* (i.e., multicore). Use the `workers` argument in `plan()` to specify how many cores you want to use. Next, run the `powRICLPM` analysis, and the power analysis will run on the specified number of cores. This can result in a significant reduction of computing time. For more information on other parallel execution strategies, see `?furrr::plan()`. 

### Progress bar using `progressr`
It can be useful to get an approximation of the progress of the `powRICLPM` analysis while running the code, especially when running the analysis in parallel. `powRICLPM()` has implemented progress notifications using the `progressr` package. Simply put, there are two options through which you can get progress notifications: 

- You can subscribe to progress updates from a specific expression by wrapping this expression with `with_progress({...})`.
- You can subscribe to progress updates from everywhere by running `handlers(global = T)`.

The second option is not fully developed yet for the `furrr` package, so instead I focus on the first. Implementing the `with_progress({...})` option, as well as parallel execution of the `powRICLPM` analysis, results in the below code:

```{r furrr-setup, eval = F}
# Load the furrr package
library(furrr)

# Check how many cores are available
future::availableCores()

# Plan the powRICLPM analysis to run on 1 core less than the number of available cores
plan(multisession, workers = 7) # For the case of 8 available cores

# Run the powRICLPM analysis 
with_progress({ # Subscribe to progress updates
  output <- powRICLPM(target_power = 0.8, # The actual power analysis function
                      search_lower = 100,
                      search_upper = 1000,
                      search_step = 50,
                      time_points = c(3, 4, 5),
                      ICC = ICC,
                      RI_cor = RI_cor, 
                      Phi = Phi,
                      within_cor = 0.3,
                      reps = 1000,
                      parameter = `wB2~wA1`)
})

# Revert back to sequential execution of code upon completion of the analysis
plan(sequential)
```

For more information about progress notification options using `progressr` for end-users, including auditory and email updates, see [https://progressr.futureverse.org](https://progressr.futureverse.org). 

## Step 6: Summarize results
The `powRICLPM()` function creates a `powRICLPM` object: A list with results, upon which we can call `print()`, `summary()`, `give()`, and `plot()` functions to print, summarize, extract results, and optionally visualize the results from the analysis, respectively.

`print()` outputs a textual summary of the power analysis design contained within the object it was called upon. It does not output any performance metrics computed by the power analysis. 

`summary()` can be used in one of four ways. First, summary can be used simply like `print()` to get information about the design of the power analysis (the different experimental conditions), as well as the number of problems the occured per condition (e.g., non-convergence, fatal estimation errors, or inadmissible results). Second, by specifying the `parameter = "..."` argument in `summary()`, the function will print the results specifically for that parameter across all experimental conditions. Third, if you specify a specific experimental condition (by specifying a sample size, number of time points and ICC using the `sample_size`, `time_points`, and `ICC` arguments), the function displays performance measures for all parameters in that experimental condition. 

```{r summary, eval = F}
# 1. General results
summary(output)

# 2. Get (slightly) more detailed parameter-specific information
summary(output, parameter = "wB2~wA1")

# 3. Display parameter-specific and condition-specific performance metrics
summary(output, sample_size = 400, time_points = 4, ICC = 0.5)

```

`give()` extracts various bits of information from an `powRICLPM` object. The exact information to be extracted is given by the `what = "..."` argument: 

1. `what = "conditions"` gives the different experimental conditions per row, where each condition is defined by a unique combination of sample size, number of time points and ICC.
2. `what = "estimation_problems"` gives the proportion of fatal errors, inadmissible values, or non-converged estimations (columns) per experimental conditions (row). 
3. `what = "results"` gives the average estimate `Avg`, minimum estimate `Min`, standard deviation of parameter estimates `stdDev`, the average standard error `SEavg`, the mean square error `MSE`, the average width of the confidence interval `Acc`, the coverage rate `Cov`, and the proportion of times the *p*-value was lower than the significance criterion `Pow`. It requires setting the `parameter = "..."` argument.
4. `what = "names"` gives the parameter names contained within the `powRICLPM` object.

```{r give, eval = F}
# 1. Extract experimental conditions
give(output, what = "conditions")

# 2. Extract estimation problems
give(output, what = "estimation_problems")

# 3. Extract results for cross-lagged effect "wB2~wA1" 
give(output, what = "results", parameter = "wB2~wA1")

# 4. Extract parameter names
give(output, what = "names")
```

Finally, `plot()` creates a `ggplot2`-plot for a specific parameter (specified using the `parameter = "..."` argument) with sample size on the x-axis, the simulated power on the y-axis, lines grouped by number of time-points, and plots wrapped by proportion of between-unit variance. `plot()` returns a `ggplot2` object that can be fully customized using `ggplot2` functionality. For example, you can change the scales, add titles, change geoms, etc. More information about options in the `ggplot2` framework can be found at [https://ggplot2-book.org/index.html](https://ggplot2-book.org/index.html). In the below example, I add a title and change the labels on the x-axis: 

```{r plot, eval = FALSE}
# Create basic plot of powRICLPM object
p <- plot(output, parameter = "wB2~wA1")
p

# Adjust plot aesthetics
p2 <- p + 
  labs(title = "Power analysis for RI-CLPM",
       caption = "Based on 1000 replications.") +
  scale_x_continuous(name = "Sample size",
                     breaks = seq(100, 1000, 100),
                     guide = guide_axis(n.dodge = 2))
p2
```






